% !TEX root = ..\main.tex
\chapter{Experiments}\label{ch:Experiments}

\section{Experiment 1: Comparison}\label{sec:Exp1}
The goal of this experiment is to provide validation of the implementation in the PoC. This is done with the results of the paper \textit{"A Comprehensive Study of Bloated Dependencies in the Maven Ecosystem"} by Soto-Valero et al. \cite{soto2020comprehensive}. In this work, a byte-code analysis is performed on Maven Artifacts to detect which of the dependencies of this artifacts are \textit{bloated}, which we refer to as \textit{unused}. Although the PoC created in this thesis does not perform exactly the same kind of analysis, we consider that a dependency is \textit{unused} if no usage is detected by any of the metrics in the model.

\subsection{Experimental set up}
Soto-Valero et al. perform a qualitative analysis, in which they analyze 31 libraries available as Maven artifacts. If unused dependencies are found during the analysis, a \textit{Pull Request} is made to the \textit{GitHub} repository of the artifact in which the unused dependencies are deleted from the \textit{pom} file.

With the information in the paper, we collect the \textit{GroupId} and \textit{ArtifactId} of the 31 artifacts. Out of the 31, 2 could not be found in the \textit{Maven Repository Central}. For the other 29, the \textit{version} to use in the experiment is determined by finding the last version released before the experiment by Soto-Valero et al. was conducted - November of 2019.

Of the 29 artifacts, 13 could not be used with the PoC, because either the artifact itself or some dependency could not be downloaded from \textit{Maven Central}, leaving a set of 16 libraries to analyse and compare the results with the results obtained by Soto-Valero et al. The table containing the identifiers of the artifacts used in this experiment can be found in Table \ref{table:comparison-artifacts}.

\begin{table}[h]
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
    Group Id              & Artifact Id                     & Version       \\
    \hline
    org.mybatis           &	mybatis	                        & 3.5.3         \\
    org.apache.flink      & flink-core                      & 1.9.1         \\
    com.puppycrawl.tools  & checkstyle                      & 8.27          \\
    com.google.auto       & auto-common                     & 0.10          \\
    edu.stanford.nlp      & stanford-corenlp                & 3.9.2         \\
    com.squareup.moshi    & moshi-kotlin                    & 1.9.2         \\
    org.neo4j             & neo4j-collections               & 3.5.13        \\
    org.asynchttpclient   & async-http-client               & 2.10.4        \\
    org.alluxio           & alluxio-core-transport          & 2.1.0         \\
    com.github.javaparser & javaparser-symbol-solver-logic  & 3.15.5        \\
    io.undertow           & undertow-benchmarks             & 2.0.27.Final  \\
    org.teavm             & teavm-core                      & 0.6.1         \\
    com.github.jknack     & handlebars-markdown             & 4.1.2         \\
    ma.glasnost.orika     & orika-eclipse-tools             & 1.5.4         \\
    fr.inria.gforge.spoon & spoon-core                      & 8.0.0         \\
    org.jacop             & jacop                           & 4.7.0         \\
    \hline
    \end{tabular}
    \end{center}
    \caption{Identifiers of the Maven artifacts used for comparison}
    \label{table:comparison-artifacts}
\end{table}

The first idea was to create a request that computed the comparison automatically. However, finding the reason why the results were not exactly the same, in most of the cases needed manual checking and research. That is why, the analysis of the libraries in Table \ref{table:comparison-artifacts} have been executed one by one, and the comparison has been done manually.

\subsection{Results}

- how many libraries have different results and why.

- overall evaluation

\section{Experiment 2: Relevance of the coupling metrics}

The goal of this experiment is to validate if the coupling metrics designed in the model, namely \texttt{MIC}, \texttt{AC}, \texttt{TMIC}, and \texttt{TAC}, are a good indicator of the usage of the dependencies by the clients. As a partial validation of the relevance of the metrics, we compare it with the results gathered from the usage metrics. We want to know how often it happens, that a dependency is used, either by using classes or methods, and it is detected as uncoupled by the coupling metrics. This way, we know if there are many cases in which a dependency is only used with a type of connection other than method invocation and field declaration.

\unsure{I'm not sure if this is the right place for the explanation of the failed data recolection, or if it is too long}

The original idea was to measure real-world data about how the clients update the dependencies and their impact on the code. We could then have seen the correlation of this impact with the degree of dependency measured with the coupling metrics. Different approaches were taken to obtain real-world data.

First, we tried to find in GitHub commits in which there had been an update of a dependency. However, the search engine in GitHub does not allow to filter the results by the language of the commit. Therefore, most of the results obtained were not useful. Also, most of the updates are only patches, which require only a bump in the version number of the declared dependency.

Based on these findings, the second approach we took was to look for updates that contained breaking changes. To find the libraries that had these type of changes, and in which versions, we used the \textit{Maven Dependency Dataset} \cite{Raemaekers2013}. Raemaekers et al. used this dataset to analyze the use of semantic versioning and the possible impact of breaking changes \cite{Raemaekers2017}. It is possible to query this dataset to obtain libraries with breaking changes, with version numbers and other libraries that depended on these. However, we need to find the commit of the client library in which the update containing a breaking change was made, and it is not always possible. We considered some of the requirements to be able to analyze a dependency with the PoC. For instance, we need all the dependencies of the client library available in Maven, and testing dependencies cannot be used since they are not analyzed by the tool. Considering all these requirements, it was not possible to obtain enough data for the experiment from the \textit{Maven Dependency Dataset} on time, since all these checks had to be done manually.

Next, we contacted the first author of the paper \textit{"Why and How Java Developers Break APIs"} \cite{Brito2018}, which mines GitHub repositories to find possible breaking changes in APIs, to obtain the dataset of breaking changes created based on their findings. Brito, the author, shared the dataset with us. The dataset includes 24 commits containing breaking changes, which correspond to 19 different libraries. Out of the 25 commits, 12 are from libraries managed by Gradle instead of Maven and cannot be used with the PoC. Besides, we could not find 4 of the commits in GitHub, and 2 others correspond to testing libraries, which are out of the scope of the analysis performed by the PoC. Therefore, there were only 6 breaking changes left, for which 3 the Maven artifact that these belong to had no dependants for which to do the analysis. The last 3 have only one dependant, and therefore is not possible to compare the impact of the breaking changes.

Finally, we tried to manually search for deprecated libraries and other libraries that used them — however, similar problems where encountered. Finding commits which replaced a deprecated dependency and the client library and all the dependencies are available in \textit{Maven Central Repository}, is a manual task that, after multiple hours of work, gave no results.


\subsection{Experimental set up}
To run this experiment, we prepared a new request in the API of the PoC. The request has to contain a path to a txt file (tab delimited). The file has to contain three columns (with headers): Group Id, Artifact Id, and version. For each one of the rows, the analysis of the dependencies is performed. The result of each of the analysis is processed, summarizing all the analysis with the following information:

\begin{itemize}
  \item \textbf{Total number of dependencies:} Number of dependencies of all the analyzed client libraries, including both, direct and indirect.
  \item \textbf{Times coupling metrics were not enough:} Number of dependencies for which all the coupling metrics had value zero, but there where methods and classes found reachable by the usage metrics.
  \item \textbf{Times MIC/TMIC were not enough:} Number of times in which there was usage found, but MIC (or TMIC in the case of transitive dependencies) had value zero.
  \item \textbf{Times AC/TAC were not enough:} Number of times in which there was usage found, but MIC (or TMIC in the case of transitive dependencies) had value zero.
  \item \textbf{List server libraries coupling metrics not enough:} The list of \textit{GroupId}, \textit{ArtifactId}, and \textit{version} of the server libraries for which all the coupling metrics were not enough to indicate if there is usage or not.
  \item \textbf{List server libraries MIC/TMIC not enough:} The list of \textit{GroupId}, \textit{ArtifactId}, and \textit{version} of the server libraries for which the metrics \texttt{MIC} and \texttt{TMIC} were not enough to indicate if there is usage or not.
  \item \textbf{List server libraries AC/TAC not enough:} The list of \textit{GroupId}, \textit{ArtifactId}, and \textit{version} of the server libraries for which the metrics \texttt{AC} and \texttt{TAC} were not enough to indicate if there is usage or not.
\end{itemize}

As can be seen, in addition to the number of times that a dependency was used and it was not detected by the metrics (or at least by one of them), the list of server libraries of these dependencies is also stored. This way, it is possible to analyze which types of libraries are those, and why the coupling metrics are not relevant in this case.

\blankl
The experiment was run with a file containing X\todo{Add number} client libraries from the \textit{Maven Central Repository}.  We selected the client libraries to use for this experiment with the following criteria. First, we used the same libraries as in the comparison experiment (see Section \ref{sec:Exp1}), but using the last version of each library. We decided to reuse these libraries because the criteria used to select these libraries by Soto-Valero et al. \cite{soto2020comprehensive} is aligned with the needs of this experiment, and are listed below:

\begin{itemize}
  \item The library is relevant - has more than 100 stars on GitHub.
  \item The library can be built successfully with Maven.
  \item Has been developed recently - in the case of Soto-Valero et al. at least October 2019.
  \item The library has at least one dependency declared.
  \item It is indicated how to create a pull request.
\end{itemize}

Although some of the items of the list are not explicitly required for our experiment. We need that the library can be built and is available in Maven as well as that it has at least one relevant dependency (compile scope). Therefore, the libraries in this set are a good fit for the experiment.

To analyze more client libraries and, therefore, more dependencies, we extended the list of libraries, according to the following process:
\todo{Finish explaining the process}

\unsure{Should I show the exact content of the file in a Table here or in an appendix at the end of the document?}

\subsection{Results}

\section{Experiment 3: Benchmarking}

\subsection{Experimental set up}

\subsection{Results}

\section{Experiment 4: Interviews}


\subsection{Experimental set up}


- Explain which stage of the visualization design is being validated, and why is domain expert interview a good idea.

- Describe the different type of questions (topic, and open question or scaled...)

- Explain the set up of the interviews (online, control to the interviewee when using the tool...)

\subsection{Results}
