% !TEX root = ../main.tex
\chapter*{Acknowledgements}
Firstly, I want to thank my family and friends for their support during all my studies, including the experience of this master thesis. In particular, to Jakob Löhnertz, for the discussions we have had about the topic, which gave me ideas to continue my work. Also, to Mar Badias, for always being there to listen to me when I needed it, working side-by-side during quarantine weeks really helped me.

I want to thank my supervisor from the University of Amsterdam, Ana Oprescu, and my supervisors within SIG Lodewijk Bergmans and Miroslav \v{Z}ivkovi\'{c}. The weekly discussions we have had during the entire process and their feedback were crucial for the success of this thesis.

To the entire research team of SIG, thank you for having me during these last months, for the help offered during the weekly standups, and for the feedback received after my presentation.

Many thanks to all the participants in the interviews about the tool I developed during the thesis. The feedback and discussions during the interviews were truly interesting and provided valuable insights.

Finally, I would like to thank the members of the organization of the Seminar Series on Advanced Techniques \& Tools for Software Evolution 2020, who allowed me to participate in the seminar. Being able to participate and organize \textit{SATToSE 2020} was a great experience. Thanks again to Ana Oprescu and Miroslav \v{Z}ivkovi\'{c} for co-authoring the paper with me; your contribution was truly valuable.
